MS Depression Paper

### 2022-08-05 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
Add files via upload

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(boot)
library(boot.pval)
#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
```

@@

  D) T values in fascicle associated with Overlap in Depression map

@@

@@

I think it works best when sticking with *volumes* rather than proportions

@@

df_good_mimosa_unique <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique)

@@

## overall
#just good mimosa
df_good_mimosa_promis <- df_people_with_promis_data_scores_added %>% filter(depGroupVar != 0)
make_demographics_table_ms_w_promis(df_good_mimosa_promis) #n=287 scans

@@

@@

@@

@@

######### functions
tval <- function(formula, data, indices) {
  d <- data[indices,] # allows boot to select sample
  fit <- lm(formula, data=d)
  return(summary(fit)$coefficients[2,3])
}

@@

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(boot)
library(boot.pval)
library(lsr)
#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
```

@@

  D) Linear models of values in fascicle associated with Overlap in Depression map

@@

    - Cohen'd correlated w/proportion of overlap of fascicle w/depression map
        - p of lm (0.001)
        -

@@

    - Cohen'd correlated w/volume of overlap of fascicle w/depression map
        - p of lm (p < 2.2 * 10^-16)
        - Largest effect in corpus callosum
I think it works best when sticking with *volumes* rather than proportions

@@

df_good_mimosa_unique <- df_demo_and_fascicles %>% 
#  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique)

@@

## overall
#just good mimosa w/promis
df_good_mimosa_promis <- df_people_with_promis_data_scores_added %>% filter(depGroupVar != 0)
make_demographics_table_ms_w_promis(df_good_mimosa_promis) #n=287 scans

@@

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
{

  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
 eval(parse(text=t_test_command))
})
names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves

@@

```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_prop_effect_size}
# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_no_cranial_nerves, function(x) 
{
  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_no_cranial_nerves

#convert to data frame
df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)

df_fascicle_dep_mask_overlap_d_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_d_values)
names(df_fascicle_dep_mask_overlap_d_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_d")

#make copy to sort
df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_d_value

df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])

# bar graph of T-values of fascicles, sorted by depression overlap
plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") + 
  geom_bar(stat="identity") +
  ylim(-0.1,0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_d_sorted_by_dep_overlap)

# plot fascicle overlap x t value
#statistic, p = 0.010
dep_overlap_x_dep_vs_healthy_d <- lm(dep_vs_healthy_d ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_d_value)
print(summary(dep_overlap_x_dep_vs_healthy_d))

scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_d_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Proportion of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis calculated in proportions (D)") + 
  ggtitle(paste0("Fascicle/depression mask Proportion overlap by dx Eff Size, p = ", round(summary(dep_overlap_x_dep_vs_healthy_d)$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_d_val_by_dep_net_overlap)

@@

```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_volume_no_CC}

#remove corpus callosum
df_fascicle_dep_mask_volume_t_value_no_cc <- df_fascicle_dep_mask_volume_t_value %>%
  filter(fascicle != "CC")

df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask <- df_fascicle_dep_mask_volume_t_value_no_cc

# bar graph of T-values of fascicles, sorted by depression overlap
plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by volume of dep network, no CC") + 
  geom_bar(stat="identity") +
  ylim(0,5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_t_sorted_by_dep_overlap)

#statistic volume, p = 0.031
dep_overlap_x_dep_vs_healthy_t_no_cc <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_t_value_no_cc)
print(summary(dep_overlap_x_dep_vs_healthy_t_no_cc))

scatter_t_val_by_dep_net_overlap_no_cc <- ggplot(data=df_fascicle_dep_mask_volume_t_value_no_cc, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis (T)") + 
  ggtitle(paste0("Fascicle/depression mask volume by dx T, no CC, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t_no_cc)$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_t_val_by_dep_net_overlap_no_cc)

```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_vol_effect_size_no_CC}
# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups

df_fascicle_and_vol_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))

###############################
#######################
### real data x dx  ###
#######################

#then, make the data frame
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_no_cranial_nerves, function(x) 
{
  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_no_cranial_nerves

#convert to data frame
df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)

#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(df_fascicle_and_vol_in_mask$fascicle, df_fascicle_and_vol_in_mask$non_zero_voxels_in_dep_map, df_d_values)
names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")

df_fascicle_dep_mask_volume_overlap_d_value_no_cc <- df_fascicle_dep_mask_volume_overlap_d_value %>%
  filter(fascicle != "CC")

#make copy to sort
df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask <- df_fascicle_dep_mask_volume_overlap_d_value_no_cc

df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$volume_in_dep_mask)])

# bar graph of eff size of fascicles, sorted by depression overlap
plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network, no CC") + 
  geom_bar(stat="identity") +
  #ylim(-0.1,0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_d_sorted_by_dep_overlap)

# plot fascicle overlap x Eff size value
#statistic, p=001
dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc )
print(summary(dep_overlap_x_dep_vs_healthy_d))

scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc , aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis calculated in Volumes (D)") + 
  ggtitle(paste0("Fascicle/depression mask Volume overlap by dx Eff Size, No CC p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_d_val_by_dep_net_overlap)

@@

######### functions
tval <- function(formula, data, indices) {
  df <- data[indices,] # allows boot to select sample
  fit <- lm(formula, data=df)
  return(summary(fit)$coefficients[2,3])
}

@@

##### effect size
# 95% confidence intervals for the correlation

set.seed(1)
data_frame_dep_v_nondep_d_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_d_value, select = c("depression_mask_overlap","dep_vs_healthy_d"))
results_prop <- boot(data=data_frame_dep_v_nondep_d_x_prop_dep_mask_overlap, statistic=tval,
   R=10000, formula=dep_vs_healthy_d~depression_mask_overlap)
print(results_prop) #actual 2.62, CI 0.589, 4.669)
plot(results_prop)
boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="perc", index=1)
print(boot_results_prop.ci)
pval_prop<-boot.pval(boot_res=results_prop, theta_null = 0, type = "perc", pval_precision = NULL)
print(pval_prop) #p=0.0083

@@

##### volume
#effect sizes
# 95% confidence intervals for the correlation
set.seed(1)
data_frame_dep_v_nondep_d_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_d_value, select = c("dep_vs_healthy_d", "volume_in_dep_mask"))

results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask, statistic=tval,
   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
print(boot_results_vol.ci)
pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
print(pval_vol) #p=1*10^-16

### with no corpus callosumm p = 1* 10^-16, CI(1.727, 5.363)
# 95% confidence intervals for the correlation
set.seed(1)
results_vol <- boot(data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc, statistic=tval,
   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
print(boot_results_vol.ci)
pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
print(pval_vol) #p=1*10^-16

@@

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})
```
